### 2024-06-25 manuscript/tables/confidence_estimate_abundance.tex

% latex table generated in R 4.4.0 by xtable 1.8-4 package
% Tue Jun 25 10:25:29 2024
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{longtable}{llll}
\caption{Confidence level on the various estimates of annual or mean abundance of vertebrate species on Bylot Island.} \\ 
  \hline
Species & Estimate & Confidence & Reason \\ 
  \hline
king eider & mean & moderate & Intensive, but opportunistic nest monitoring (8 km2) \\ 
  long-tailed duck & mean & moderate & Intensive, but opportunistic nest monitoring (8 km2) \\ 
  Pacific loon & 2017-2019, 2022 & high & Intensive study area-wide nest monitoring (389 km2) \\ 
  red-throated loon & 2017-2019, 2022 & high & Intensive study area-wide nest monitoring (389 km2) \\ 
  cackling goose & 2004-2016 & moderate & \makecell[tl]{Extrapolation from exponential model of growth (R2=0.74,\\p=0.15, n=5)} \\ 
  cackling goose & 2017-2019, 2022-2023 & high & Intensive study area-wide nest monitoring (389 km2) \\ 
  snow goose & 1999-2009 & low & \makecell[tl]{Density based only on nest monitoring and uncertainty\\on goose colony area (mean)} \\ 
  snow goose & 2010-2019, 2022-2023 & high & \makecell[tl]{Intensive study area-wide monitoring based on several\\methods and annual colony outline} \\ 
  tundra swan & 2017-2019, 2022 & high & Intensive study area-wide nest monitoring (389 km2) \\ 
  rough-legged hawk & 2007-2012 & high & \makecell[tl]{Extrapolation from intensive nest monitoring (111 km2,\\R2=0.99, p<0.0001, n=8)} \\ 
  rough-legged hawk & 2013-2019, 2022 & high & Intensive study area-wide nest monitoring (389 km2) \\ 
  peregrine falcon & 2013-2019, 2022 & high & Intensive study area-wide nest monitoring (389 km2) \\ 
  snowy owl & 1996-2011 & high & \makecell[tl]{Extrapolation from intensive nest monitoring (111 km2,\\R2=0.99, p<0.0001, n=10)} \\ 
  snowy owl & 2012-2019, 2022-2023 & high & Intensive study area-wide nest monitoring (389 km2) \\ 
  rock ptarmigan & mean & low & \makecell[tl]{Intensive, but opportunistic nest monitoring (8 km2)\\and prime nesting habitat not well sampled} \\ 
  sandhill crane & mean & moderate & \makecell[tl]{Extrapolation from intensive nest monitoring (8 km2)\\and uncertain relation with large scale indices} \\ 
  American golden-plover & 2014-2019, 2022-2023 & high & Distance sampling throughout lowland (313 km2) \\ 
  black-bellied plover & mean & moderate & Derived from the abundance estimate of another species \\ 
  common-ringed plover & 2015-2017 & moderate & Nest monitoring on the main breeding sites \\ 
  ruddy turnstone & mean & low & Derived from the abundance estimate of another species \\ 
  red knot & mean & low & Derived from the abundance estimate of another species \\ 
  pectoral sandpiper & mean & low & Derived from the abundance estimate of another species \\ 
  Baird's sandpiper & mean & moderate & \makecell[tl]{Extrapolation from intensive nest monitoring (2 km2)\\and uncertain relation with large scale indices} \\ 
  white-rumped sandpiper & mean & low & Derived from the abundance estimate of another species \\ 
  buff-breasted sandpiper & mean & low & Derived from the abundance estimate of another species \\ 
  red phalarope & mean & low & Derived from the abundance estimate of another species \\ 
  glaucous gull & 2004-2016 & high & \makecell[tl]{Extrapolation from intensive nest monitoring (111 km2,\\R2=0.84, p=0.16, n=4)} \\ 
  glaucous gull & 2017-2019, 2022 & high & Intensive study area-wide nest monitoring (389 km2) \\ 
  long-tailed jaeger & 2004-2019, 2022 & high & \makecell[tl]{Extrapolation based on intensive nest monitoring (33\\km2)} \\ 
  parasitic jaeger & mean & low & \makecell[tl]{Opportunistic nest monitoring and uncertainty about number\\of non-breeding individuals } \\ 
  common raven & mean & low & \makecell[tl]{Low density and uncertainty about number of non-breeding\\individuals} \\ 
  horned lark & mean & moderate & Derived from the abundance estimate of another species \\ 
  American pipit & mean & low & \makecell[tl]{Derived from the abundance estimate of another species\\and prime nesting habitat not well sampled} \\ 
  Lapland longspur & mean & moderate & \makecell[tl]{Extrapolation from intensive nest monitoring (2 km2)\\and uncertain relation with large scale indices} \\ 
  snow bunting & mean & low & \makecell[tl]{Derived from the abundance estimate of another species\\and prime nesting habitat not well sampled} \\ 
  brown lemming & 1995-2019, 2021-2022 & moderate & \makecell[tl]{Rigorous density estimates but at a small spatial scale\\(0.22 km2)} \\ 
  collared lemming & 1995-2019, 2021-2022 & moderate & \makecell[tl]{Rigorous density estimates but at a small spatial scale\\(0.22 km2)} \\ 
  Arctic hare & mean & low & Prime habitat not well sampled \\ 
  ermine & 1993-2019 & moderate & \makecell[tl]{Based on indirect indices and uncertainty on ermine home\\range size estimates} \\ 
  Arctic fox & mean & moderate & Derived from extensive fox home range size studies (n=109) \\ 
   \hline
\hline
\label{table:confidence_estimate_abundance}
\end{longtable}
\endgroup