### 2024-06-25 manuscript/tables/interannual_nest_density_variation.tex

% latex table generated in R 4.4.0 by xtable 1.8-4 package
% Tue Jun 25 10:25:27 2024
\begin{table}[ht]
\centering
\caption{Annual nest density (nests/km2) of selected avian species estimated on different zones of Bylot Island.} 
\label{table:interannual_nest_density_variation}
\begingroup\fontsize{10pt}{10pt}\selectfont
\begin{tabularx}{0.9\textwidth}{lllr}
  \hline
Species & Zone & Mean ± SD & Number of years \\ 
  \hline
Baird's sandpiper & qarlikturvik (2x1 km plot) & 5.000 ± 3.558 &  17 \\ 
  Lapland longspur & qarlikturvik (2x1 km plot) & 13.559 ± 5.849 &  17 \\ 
  king eider & qarlikturvik (4x2 km plot) & 0.115 ± 0.138 &  16 \\ 
  long-tailed duck & qarlikturvik (4x2 km plot) & 0.092 ± 0.138 &  16 \\ 
  rock ptarmigan & qarlikturvik (4x2 km plot) & 0.031 ± 0.055 &  16 \\ 
  sandhill crane & qarlikturvik (4x2 km plot) & 0.161 ± 0.087 &  16 \\ 
  long-tailed jaeger & qarlikturvik valley & 0.362 ± 0.380 &  17 \\ 
  Pacific loon & whole study area & 0.005 ± 0.004 &   4 \\ 
  red-throated loon & whole study area & 0.082 ± 0.019 &   4 \\ 
  cackling goose & whole study area & 0.177 ± 0.064 &   5 \\ 
  tundra swan & whole study area & 0.001 ± 0.001 &   4 \\ 
  rough-legged hawk & whole study area & 0.157 ± 0.151 &   8 \\ 
  peregrine falcon & whole study area & 0.053 ± 0.007 &   8 \\ 
  snowy owl & whole study area & 0.022 ± 0.058 &  10 \\ 
  common ringed plover & whole study area & 0.070 ± 0.012 &   3 \\ 
  glaucous gull & whole study area & 0.091 ± 0.011 &   4 \\ 
  parasitic jaeger & whole study area & 0.010 ± 0.004 &  12 \\ 
  common raven & whole study area & 0.003 ± 0.003 &   8 \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}