### 2024-06-25 manuscript/tables/species_mean_abudance.tex

% latex table generated in R 4.4.0 by xtable 1.8-4 package
% Tue Jun 25 10:25:29 2024
\begin{table}[ht]
\centering
\caption{Mean (± standard deviation) and minimum and maximum (when available) annual abundance of vertebrate species for the whole Bylot Island study area (389 km2). In some cases, several independent approaches have been used to estimate the abundance of a species as a proxy for uncertainty.} 
\label{table:species_mean_abudance}
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{tabularx}{\textwidth}{lllrrr}
  \hline
Species & Method & Mean ± sd & n & Minimum & Maximum \\ 
  \hline
king eider & nest sampling (extrapolation habitat) & 25 &  &  &  \\ 
  king eider & incidental observations (relative abundance) & 106 &  &  &  \\ 
  long-tailed duck & nest sampling (extrapolation habitat) & 20 &  &  &  \\ 
  long-tailed duck & incidental observations (relative abundance) & 191 &  &  &  \\ 
  Pacific loon & nest sampling & 4 ± 3 &   4 &   0 &   6 \\ 
  red-throated loon & nest sampling & 64 ± 15 &   4 &  42 &  76 \\ 
  cackling goose & nest sampling & 50 ± 59 &  28 &   2 & 214 \\ 
  snow goose & combined methods & 25289 ± 13327 &  23 & 2505 & 49076 \\ 
  tundra swan & nest sampling & 1 ± 1 &   4 &   0 &   2 \\ 
  rough-legged hawk & nest sampling & 25 ± 24 &  16 &   0 &  66 \\ 
  peregrine falcon & nest sampling & 10 ± 1 &   8 &   8 &  12 \\ 
  snowy owl & nest sampling & 16 ± 33 &  26 &   0 & 144 \\ 
  rock ptarmigan & nest sampling (extrapolation habitat) & 24 &  &  &  \\ 
  sandhill crane & nest sampling (extrapolation transects) & 34 &  &  &  \\ 
  American golden-plover & distance sampling & 774 ± 275 &   8 & 306 & 1040 \\ 
  black-bellied plover & transects (relative abundance) & 22 &  &  &  \\ 
  black-bellied plover & incidental observations (relative abundance) & 61 &  &  &  \\ 
  common-ringed plover & nest sampling & 55 ± 9 &   3 &  44 &  62 \\ 
  ruddy turnstone & transects (relative abundance) & 40 &  &  &  \\ 
  ruddy turnstone & incidental observations (relative abundance) & 54 &  &  &  \\ 
  red knot & transects (relative abundance) & 64 &  &  &  \\ 
  red knot & incidental observations (relative abundance) & 237 &  &  &  \\ 
  pectoral sandpiper & transects (relative abundance) & 79 &  &  &  \\ 
  pectoral sandpiper & incidental observations (relative abundance) & 260 &  &  &  \\ 
  Baird's sandpiper & nest sampling (extrapolation transects) & 2495 &  &  &  \\ 
  white-rumped sandpiper & transects (relative abundance) & 1135 &  &  &  \\ 
  white-rumped sandpiper & incidental observations (relative abundance) & 1156 &  &  &  \\ 
  buff-breasted sandpiper & transects (relative abundance) & 41 &  &  &  \\ 
  buff-breasted sandpiper & incidental observations (relative abundance) & 8 &  &  &  \\ 
  red phalarope & transects (relative abundance) & 156 &  &  &  \\ 
  red phalarope & incidental observations (relative abundance) & 276 &  &  &  \\ 
  glaucous gull & nest sampling & 73 ± 7 &  17 &  59 &  80 \\ 
  long-tailed jaeger & nest sampling & 272 ± 285 &  17 &   0 & 900 \\ 
  parasitic jaeger & banding & 17 &  &  &  \\ 
  parasitic jaeger & nest sampling & 8 &  &  &  \\ 
  common raven & transects (relative abundance) & 14 &  &  &  \\ 
  common raven & incidental observations (relative abundance) & 31 &  &  &  \\ 
  horned lark & transects (relative abundance) & 362 &  &  &  \\ 
  horned lark & incidental observations (relative abundance) & 411 &  &  &  \\ 
  American pipit & transects (relative abundance) & 53 &  &  &  \\ 
  American pipit & incidental observations (relative abundance) & 87 &  &  &  \\ 
  Lapland longspur & nest sampling (extrapolation transects) & 7110 &  &  &  \\ 
  snow bunting & transects (relative abundance) & 18 &  &  &  \\ 
  snow bunting & incidental observations (relative abundance) & 276 &  &  &  \\ 
  brown lemming & trapping & 54043 ± 93530 &  27 &   0 & 447630 \\ 
  collared lemming & trapping & 8128 ± 10334 &  27 &   0 & 39302 \\ 
  Arctic hare & incidental observations (relative abundance) & 6 &  &  &  \\ 
  ermine & testimonials & 40 ± 37 &  27 &   8 & 156 \\ 
  Arctic fox & individual home range & 53 &  &  &  \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}