### 2024-06-25 manuscript/tables/species_name_strategy.tex

% latex table generated in R 4.4.0 by xtable 1.8-4 package
% Tue Jun 25 10:25:25 2024
\begin{table}[ht]
\centering
\caption{Species composition of the vertebrate community of Bylot Island with the corresponding annual cycle strategy (i.e., resident, partial migrant, migrant).} 
\label{table:species_name_strategy}
\begingroup\fontsize{10pt}{10pt}\selectfont
\begin{tabularx}{0.95\textwidth}{llll}
  \hline
Functional group & Scientific name & Name & Annual cycle strategy \\ 
  \hline
ducks and loons  & Somateria spectabilis & king eider & migrant \\ 
  ducks and loons  & Clangula hyemalis & long-tailed duck & migrant \\ 
  ducks and loons  & Gavia pacifica & Pacific loon & migrant \\ 
  ducks and loons  & Gavia stellata & red-throated loon & migrant \\ 
  geese and swans  & Branta hutchinsii & cackling goose & migrant \\ 
  geese and swans  & Anser caerulescens & snow goose & migrant \\ 
  geese and swans  & Cygnus columbianus & tundra swan & migrant \\ 
  raptors  & Buteo lagopus & rough-legged hawk & migrant \\ 
  raptors  & Falco peregrinus & peregrine falcon & migrant \\ 
  raptors  & Bubo scandiacus & snowy owl & migrant \\ 
  ptarmigans  & Lagopus muta & rock ptarmigan & resident \\ 
  cranes  & Antigone canadensis & sandhill crane & migrant \\ 
  shorebirds  & Pluvialis dominica & American golden-plover & migrant \\ 
  shorebirds  & Pluvialis squatarola & black-bellied plover & migrant \\ 
  shorebirds  & Charadrius hiaticula & common-ringed plover & migrant \\ 
  shorebirds  & Arenaria interpres & ruddy turnstone & migrant \\ 
  shorebirds  & Calidris canutus & red knot & migrant \\ 
  shorebirds  & Calidris melanotos & pectoral sandpiper & migrant \\ 
  shorebirds  & Calidris bairdii & Baird's sandpiper & migrant \\ 
  shorebirds  & Calidris fuscicollis & white-rumped sandpiper & migrant \\ 
  shorebirds  & Calidris subruficollis & buff-breasted sandpiper & migrant \\ 
  shorebirds  & Phalaropus fulicarius & red phalarope & migrant \\ 
  gulls  & Larus hyperboreus & glaucous gull & migrant \\ 
  jaegers  & Stercorarius longicaudus & long-tailed jaeger & migrant \\ 
  jaegers  & Stercorarius parasiticus & parasitic jaeger & migrant \\ 
  ravens  & Corvus corax & common raven & partial migrant \\ 
  passerines  & Eremophila alpestris & horned lark & migrant \\ 
  passerines  & Anthus rubescens & American pipit & migrant \\ 
  passerines  & Calcarius lapponicus & Lapland longspur & migrant \\ 
  passerines  & Plectrophenax nivalis & snow bunting & migrant \\ 
  lemmings  & Lemmus trimucronatus & brown lemming & resident \\ 
  lemmings  & Dicrostonyx groenlandicus & collared lemming & resident \\ 
  hares  & Lepus arcticus & Arctic hare & resident \\ 
  ermines  & Mustela richardsonii & ermine & resident \\ 
  foxes  & Vulpes lagopus & Arctic fox & partial migrant \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}