### 2024-06-25 manuscript/tables/species_year_monitoring.tex

% latex table generated in R 4.4.0 by xtable 1.8-4 package
% Tue Jun 25 10:25:26 2024
\begin{table}[ht]
\centering
\caption{Summary of vertebrate species monitoring in the Bylot Island study area. In this paper, we excluded certain years for specific species due to reduced sampling efforts. As a result, duration of times series presented here may differ slightly from those in in Gauthier et al. (2024b).} 
\label{table:species_year_monitoring}
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{tabularx}{\textwidth}{lllll}
  \hline
Species & Zone & Years & Monitoring & References \\ 
  \hline
snow goose & camp 2 & 1999-2019, 2023 & systematic & Gauthier et Cadieux, 2020 \\ 
  rough-legged hawk & qarlikt., black \& south plat. & 2007-2019, 2022 & systematic & Gauthier et al., 2020 \\ 
  peregrine falcon & qarlikt., black \& south plat. & 2007-2019, 2022 & systematic & Gauthier et al., 2020 \\ 
  snowy owl & qarlikt., black \& south plat. & 1996-2019, 2023 & systematic & Gauthier et al., 2020 \\ 
  Baird's sandpiper & qarlikturvik (2x1 km plot) & 2005-2019, 2022-2023 & systematic & Bêty, 2020 \\ 
  Lapland longspur & qarlikturvik (2x1 km plot) & 2005-2019, 2022-2023 & systematic & Gauthier and Bêty, 2020 \\ 
  king eider & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & opportunistic & unpublished data \\ 
  long-tailed duck & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & opportunistic & unpublished data \\ 
  rock ptarmigan & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & opportunistic & unpublished data \\ 
  sandhill crane & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & opportunistic & unpublished data \\ 
  brown lemming & qarlikturvik (trapping grids) & 1995-2019, 2021-2022 & systematic & Gauthier, 2020 \\ 
  collared lemming & qarlikturvik (trapping grids) & 1995-2019, 2021-2022 & systematic & Gauthier, 2020 \\ 
  Pacific loon & qarlikturvik valley & 2004-2019, 2022 & systematic & unpublished data \\ 
  red-throated loon & qarlikturvik valley & 2004-2019, 2022 & systematic & unpublished data \\ 
  cackling goose & qarlikturvik valley & 2004-2019, 2022-2023 & systematic & unpublished data \\ 
  tundra swan & qarlikturvik valley & 2004-2019, 2022 & systematic & unpublished data \\ 
  glaucous gull & qarlikturvik valley & 2004-2019, 2022 & systematic & Gauthier et al., 2020 \\ 
  long-tailed jaeger & qarlikturvik valley & 2004-2019, 2022 & systematic & Gauthier et al., 2020 \\ 
  parasitic jaeger & qarlikturvik valley & 2004-2019, 2022 & systematic & unpublished data \\ 
  Pacific loon & whole study area & 2017-2019, 2022 & systematic & Duchesne et al., 2021 \\ 
  red-throated loon & whole study area & 2017-2019, 2022 & systematic & Duchesne et al., 2021 \\ 
  cackling goose & whole study area & 2017-2019, 2022-2023 & systematic & Duchesne et al., 2021 \\ 
  tundra swan & whole study area & 2017-2019, 2022 & systematic & Duchesne et al., 2021 \\ 
  rough-legged hawk & whole study area & 2013-2019, 2022 & systematic & Gauthier et al., 2020 \\ 
  peregrine falcon & whole study area & 2013-2019, 2022  & systematic & Gauthier et al., 2020 \\ 
  snowy owl & whole study area & 2012-2019, 2022-2023 & systematic & Gauthier et al., 2020 \\ 
  common-ringed plover & whole study area & 2015-2017 & systematic & Bêty, 2020 \\ 
  glaucous gull & whole study area & 2017-2019, 2022 & systematic & Gauthier et al., 2020 \\ 
  parasitic jaeger & whole study area & 2009-2019, 2022 & opportunistic & unpublished data \\ 
  common raven & whole study area & 2013-2019, 2022 & systematic & unpublished data \\ 
  ermine & whole study area & 1993-2019 & opportunistic & Bolduc et al., 2023 \\ 
  Arctic fox & whole study area & 2008-2016 & systematic & Dulude-de Broin et al., 2023 \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}