### 2024-06-25 manuscript/tables/table_summary_methods.tex

% latex table generated in R 4.4.0 by xtable 1.8-4 package
% Tue Jun 25 10:25:27 2024
\begin{table}[H]
\centering
\caption{Summary of the methods and spatial scales used to estimate the density of each vertebrate species on Bylot Island, and how abundance was extrapoated at the landscape scale.} 
\label{table:summary_methods}
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{tabularx}{\textwidth}{lllll}
  \hline
Species & Sampling method & Reference species & Sampling scale & Spatial extrapolation \\ 
  \hline
king eider & nest sampling &  & 4x2km plot & wetlands \\ 
  long tailed duck & nest sampling &  & 4x2km plot & wetlands \\ 
  king eider & relat. abund. (incidental obs.) & red-throated loon & study area &  \\ 
  long tailed duck & relat. abund. (incidental obs.) & red-throated loon & study area &  \\ 
  Pacific loon & nest sampling &  & study area &  \\ 
  red-throated loon & nest sampling &  & study area &  \\ 
  cackling goose & nest sampling &  & study area &  \\ 
  snow goose & nest sampling &  & camp 2 & goose colony \\ 
  snow goose & vertebrate count transects &  & lowland & goose colony \\ 
  snow goose & snow goose point counts &  & lowland & goose colony \\ 
  tundra swan & nest sampling &  & study area &  \\ 
  rough-legged hawk & nest sampling &  & study area &  \\ 
  peregrine falcon & nest sampling &  & study area &  \\ 
  snowy owl & nest sampling &  & study area &  \\ 
  rock ptarmigan & nest sampling &  & 4x2km plot & all habitats \\ 
  sandhill crane & nest sampling &  & 4x2km plot & transects \\ 
  American golden-plover & distance sampling &  & lowland & transects \\ 
  black-bellied plover & relat. abund. (transects) & A. golden-plover & lowland & transects \\ 
  black-bellied plover & relat. abund. (incidental obs.) & A. golden-plover & study area &  \\ 
  common-ringed plover & nest sampling &  & study area &  \\ 
  ruddy turnstone & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
  ruddy turnstone & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
  red knot & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
  red knot & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
  pectoral sandpiper & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
  pectoral sandpiper & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
  Baird’s sandpiper & nest sampling &  & 2x1km plot & transects \\ 
  white-rumped sandpiper & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
  white-rumped sandpiper & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
  buff-breasted sandpiper & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
  buff-breasted sandpiper & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
  red phalarope & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
  red phalarope & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
  glaucous gull & nest sampling &  & study area &  \\ 
  long-tailed jaeger & nest sampling &  & qarlikturvik & mesic habitat \\ 
  parasitic jaeger & maximum banded individual &  & study area &  \\ 
  parasitic jaeger & nest sampling &  & study area &  \\ 
  common raven & nest sampling &  & study area &  \\ 
  common raven & relat. abund. (incidental obs.) & Parasitic jaeger & study area &  \\ 
  horned lark & relat. abund. (transects) & lapland longspur & lowland & transects \\ 
  horned lark & relat. abund. (incidental obs.) & lapland longspur & study area &  \\ 
  American pipit & relat. abund. (transects) & lapland longspur & lowland & transects \\ 
  American pipit & relat. abund. (incidental obs.) & lapland longspur & study area &  \\ 
  lapland longspur & nest sampling &  & 2x1km plot & transects \\ 
  snow bunting & relat. abund. (transects) & lapland longspur & lowland & transects \\ 
  snow bunting & relat. abund. (incidental obs.) & lapland longspur & study area &  \\ 
  brown lemming & capture marking recapture &  & two 11 ha grids & wet and mesic hab. \\ 
  collared lemming & capture marking recapture &  & two 11 ha grids & wet and mesic hab. \\ 
  arctic hare & incidental observations & Arctic fox & study area &  \\ 
  ermine & relat. abund. (testimonials) &  & study area &  \\ 
  arctic fox & home range (movement track.) &  & study area &  \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}