### 2024-09-18 manuscript/tables/species_mean_abudance.tex
Combine table 7 and 8

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Tue Sep 10 14:57:19 2024
\begin{table}[ht]
\centering
\caption{Selected methods to represent the mean (± standard deviation) of annual abundance of vertebrate species for the whole Bylot Island study area (389 km2).} 
\label{table:species_mean_abudance}
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{tabularx}{0.85\textwidth}{llllr}
  \hline
Species & Method & Mean & sd & n \\ 
  \hline
king eider & incidental observations (relative abundance) & 106 &  &  \\ 
  long-tailed duck & incidental observations (relative abundance) & 191 &  &  \\ 
  Pacific loon & nest sampling & 4 & 3 &   4 \\ 
  red-throated loon & nest sampling & 64 & 15 &   4 \\ 
  cackling goose & nest sampling & 140 & 41 &   7 \\ 
  snow goose & combined methods & 31852 & 12092 &  12 \\ 
  tundra swan & nest sampling & 1 & 1 &   4 \\ 
  rough-legged hawk & nest sampling & 25 & 24 &  16 \\ 
  peregrine falcon & nest sampling & 10 & 1 &   8 \\ 
  snowy owl & nest sampling & 16 & 33 &  26 \\ 
  rock ptarmigan & nest sampling (extrapolation habitat) & 24 &  &  \\ 
  sandhill crane & nest sampling (extrapolation transects) & 34 &  &  \\ 
  American golden-plover & distance sampling & 774 & 275 &   8 \\ 
  black-bellied plover & transects (relative abundance) & 22 &  &  \\ 
  common-ringed plover & nest sampling & 55 & 9 &   3 \\ 
  ruddy turnstone & transects (relative abundance) & 40 &  &  \\ 
  red knot & transects (relative abundance) & 64 &  &  \\ 
  pectoral sandpiper & transects (relative abundance) & 79 &  &  \\ 
  Baird's sandpiper & nest sampling (extrapolation transects) & 2495 &  &  \\ 
  white-rumped sandpiper & transects (relative abundance) & 1135 &  &  \\ 
  buff-breasted sandpiper & transects (relative abundance) & 41 &  &  \\ 
  red phalarope & transects (relative abundance) & 156 &  &  \\ 
  glaucous gull & nest sampling & 73 & 7 &  17 \\ 
  long-tailed jaeger & nest sampling (extrapolation habitat) & 272 & 285 &  17 \\ 
  parasitic jaeger & banding & 17 &  &  \\ 
  common raven & incidental observations (relative abundance) & 31 &  &  \\ 
  horned lark & incidental observations (relative abundance) & 411 &  &  \\ 
  American pipit & incidental observations (relative abundance) & 87 &  &  \\ 
  Lapland longspur & nest sampling (extrapolation transects) & 7110 &  &  \\ 
  snow bunting & incidental observations (relative abundance) & 276 &  &  \\ 
  brown lemming & trapping & 54043 & 93530 &  27 \\ 
  collared lemming & trapping & 8128 & 10334 &  27 \\ 
  Arctic hare & incidental observations (relative abundance) & 6 &  &  \\ 
  ermine & testimonials & 40 & 37 &  27 \\ 
  Arctic fox & individual home range & 53 &  &  \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}