### 2024-09-18 manuscript/tables/table_summary_methods.tex

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Tue Sep 10 15:05:42 2024
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{longtable}{|p{0.11\textwidth}|p{0.07\textwidth}|p{0.21\textwidth}|p{0.06\textwidth}|p{0.18\textwidth}|p{0.035\textwidth}|p{0.037\textwidth}|p{0.037\textwidth}|p{0.035\textwidth}|p{0.017\textwidth}|p{0.14\textwidth}|}
\caption{Summary of the mean (± standard deviation) and minimum and maximum (when available) annual abundance of vertebrate species for the whole Bylot Island study area (389 km2).  In some cases, several independent approaches have been used to estimate the abundance of a species as a proxy for uncertainty. We did not quantitatively assess uncertainty on each estimate of species abundance. However, we provide a qualitative measure of method quality based on the quality and quantity of data available, method used for extrapolation (if necessary), and in some cases, from the fit of statistical models to estimate density. We also provided coarse confidence intervals derived from consultation with field experts at the study site.} \\ 
  \hline
{\textbf{Species}} & {\textbf{Estimate}} & {\textbf{Method}} & {\textbf{Method quality}} & {\textbf{Justification}} & {\textbf{Min}} & {\textbf{Max}} & {\textbf{Mean}} & {\textbf{sd}} & {\textbf{n}} & {\textbf{Expert based confidence interval}} \\ 
  \hline
snow goose & 1999-2009 & Nest monitoring plots extrapolated to mean goose colony area & low & Relatively small sample size and uncertainty on goose colony area  & 2505 & 35404 & 18129 & 11037 &  11 & several thousands to few tens of thousands \\ 
   \hline
snow goose & 2010-2019, 2022-2023 & Intensive study area-wide monitoring based on several
methods and annual colony outline & high & Multiple independent methods and annual colony outline & 8687 & 49076 & 31852 & 12092 &  12 & thousands \\ 
   \hline
snowy owl & 1996-2011 & Extrapolation from intensive nest monitoring (111 km2,
R2=0.99, p$<$0.0001, n=10) & high & Strong correlation with study area-wide nest density &   0 &  67 & 15 & 24 &  16 & units to few tens \\ 
   \hline
snowy owl & 2012-2019, 2022-2023 & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 & 144 & 17 & 45 &  10 & units to few tens \\ 
   \hline
glaucous gull & 2004-2016 & Extrapolation from intensive nest monitoring (111 km2,
R2=0.84, p=0.16, n=4) & high & Strong correlation with study area-wide nest density &  59 &  80 & 73 & 6 &  13 & few tens \\ 
   \hline
glaucous gull & 2017-2019, 2022 & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &  60 &  80 & 71 & 9 &   4 & few tens \\ 
   \hline
peregrine falcon & 2013-2019, 2022 & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   8 &  12 & 10 & 1 &   8 & units \\ 
   \hline
rough-legged hawk & 2007-2012 & Extrapolation from intensive nest monitoring (111 km2,
R2=0.99, p$<$0.0001, n=8) & high & Strong correlation with study area-wide nest density &   3 &  59 & 20 & 19 &   8 & few tens \\ 
   \hline
rough-legged hawk & 2013-2019, 2022 & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 &  66 & 30 & 29 &   8 & few tens \\ 
   \hline
American golden-plover & 2014-2019, 2022-2023 & Distance sampling throughout lowland (313 km2) & high & Large sample size & 306 & 1040 & 774 & 275 &   8 & hundreds \\ 
   \hline
cackling goose & 1996-2016 & Extrapolation from exponential model of growth (R2=0.74,
p=0.15, n=5) & moderate & Strong correlation with opportunistic nest monitoring &   2 & 158 & 31 & 41 &  23 & units to tens \\ 
   \hline
cackling goose & 2017-2019, 2022-2023 & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &  80 & 214 & 138 & 50 &   5 & few tens \\ 
   \hline
Arctic fox & mean & Derived from extensive fox home range size studies (n=109) & moderate & Indirect indices, but large sample size &  &  & 53 &  &  & few tens \\ 
   \hline
collared lemming & 1995-2019, 2021-2022 & Rigorous density estimates at small spatial scale
(0.22 km2) extrapolated by habitat & moderate & Intensive sampling, but small spatial coverage and extrapolation by habitat &   0 & 39302 & 8128 & 10334 &  27 & few thousands to tens of thousands \\ 
   \hline
brown lemming & 1995-2019, 2021-2022 & Rigorous density estimates at small spatial scale
(0.22 km2) extrapolated by habitat & moderate & Intensive sampling, but small spatial coverage and extrapolation by habitat &   0 & 447630 & 54043 & 93530 &  27 & few thousands to few hundreds of thousands \\ 
   \hline
ermine & 1993-2019 & Indices of relative abundance derived from testimonials converted to abundance using home range size & moderate & Indirect indices and uncertainty on ermine home range size estimates &   8 & 156 & 40 & 37 &  27 & tens \\ 
   \hline
long-tailed jaeger & 2004-2019, 2022 & Intensive nest monitoring (33 km2) extrapolated by habitat & high & Relatively large spatial coverage of sampling &   0 & 900 & 272 & 285 &  17 & units to few hundreds \\ 
   \hline
red-throated loon & 2017-2019, 2022 & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &  42 &  76 & 64 & 15 &   4 & few tens \\ 
   \hline
Pacific loon & 2017-2019, 2022 & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 &   6 & 4 & 3 &   4 & units \\ 
   \hline
tundra swan & 2017-2019, 2022 & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 &   2 & 1 & 1 &   4 & units \\ 
   \hline
common-ringed plover & 2015-2017 & Nest monitoring on the main breeding sites & moderate & Intensive monitoring, but not exhaustive to study area &  44 &  62 & 55 & 9 &   3 & few tens \\ 
   \hline
black-bellied plover & mean & Derived from the abundance estimate of American golden-plover using transects observations & moderate & Derived from high quality estimate of another species &  &  & 22 &  &  & few tens \\ 
   \hline
black-bellied plover & mean & Derived from the abundance estimate of American golden-plover using incidental observations & moderate & Derived from high quality estimate of another species &  &  & 61 &  &  & few tens \\ 
   \hline
Lapland longspur & mean & Extrapolation from intensive nest monitoring (2 km2)
and transects observations (R2=0.62, p=0.1, n=8) & moderate & Uncertain relation with large scale indices &  &  & 7110 &  &  & few thousands \\ 
   \hline
Baird's sandpiper & mean & Extrapolation from intensive nest monitoring (2 km2)
and transects observations (R2=0.38, p=0.35, n=8) & moderate & Uncertain relation with large scale indices &  &  & 2495 &  &  & several hundreds to few thousands \\ 
   \hline
sandhill crane & mean & Extrapolation from intensive nest monitoring (8 km2) and transect observations & moderate & Uncertain relation with large scale indices &  &  & 34 &  &  & few tens \\ 
   \hline
king eider & mean & Intensive, but opportunistic nest monitoring (8 km2) extrapolated by habitat & moderate & Intensive, but opportunistic monitoring at relatively small spatial scale &  &  & 25 &  &  & tens \\ 
   \hline
king eider & mean & Derived from the abundance estimate of red-throated loon using incidental observations & moderate & Derived from high quality estimate of another species &  &  & 106 &  &  & tens \\ 
   \hline
long-tailed duck & mean & Intensive, but opportunistic nest monitoring (8 km2) extrapolated by habitat & moderate & Intensive, but opportunistic monitoring at relatively small spatial scale &  &  & 20 &  &  & tens \\ 
   \hline
long-tailed duck & mean & Derived from the abundance estimate of red-throated loon using incidental observations & moderate & Derived from high quality estimate of another species &  &  & 191 &  &  & tens \\ 
   \hline
rock ptarmigan & mean & Intensive, but opportunistic nest monitoring (8 km2) extrapolated to study area & low & Intensive, but opportunistic monitoring at relatively small spatial scale and prime nesting habitat not well sampled &  &  & 24 &  &  & tens \\ 
   \hline
horned lark & mean & Derived from the abundance estimate of Lapland longspur using transects observations & low & Derived from moderate quality estimate of another species &  &  & 362 &  &  & few hundreds \\ 
   \hline
horned lark & mean & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 411 &  &  & few hundreds \\ 
   \hline
ruddy turnstone & mean & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 40 &  &  & tens \\ 
   \hline
ruddy turnstone & mean & Derived from the abundance estimate of Bairds sandpiper using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 54 &  &  & tens \\ 
   \hline
red phalarope & mean & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 156 &  &  & tens \\ 
   \hline
red phalarope & mean & Derived from the abundance estimate of Bairds sandpiper using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 276 &  &  & tens \\ 
   \hline
red knot & mean & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 64 &  &  & tens \\ 
   \hline
red knot & mean & Derived from the abundance estimate of Bairds sandpiper using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 237 &  &  & tens \\ 
   \hline
white-rumped sandpiper & mean & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 1135 &  &  & hundreds \\ 
   \hline
white-rumped sandpiper & mean & Derived from the abundance estimate of Bairds sandpiper using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 1156 &  &  & hundreds \\ 
   \hline
buff-breasted sandpiper & mean & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 41 &  &  & tens \\ 
   \hline
buff-breasted sandpiper & mean & Derived from the abundance estimate of Bairds sandpiper using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 8 &  &  & tens \\ 
   \hline
pectoral sandpiper & mean & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 79 &  &  & tens \\ 
   \hline
pectoral sandpiper & mean & Derived from the abundance estimate of Bairds sandpiper using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 260 &  &  & tens \\ 
   \hline
parasitic jaeger & mean & Maximum number of nest found annually during study area-wide opportunistic nest monitoring & low & Uncertainty about number of non-breeding individuals  &  &  & 17 &  &  & few tens \\ 
   \hline
parasitic jaeger & mean & Maximum number of individuals banded in a year & low & Based on a single year and potentially not all individuals were captured &  &  & 8 &  &  & few tens \\ 
   \hline
common raven & mean & Derived from the abundance estimate of glaucous gull using transects observations & low & Derived from moderate quality estimate of another species, but potential difference in detectability between species &  &  & 14 &  &  & few tens \\ 
   \hline
common raven & mean & Derived from the abundance estimate of glaucous gull using incidental observations & low & Derived from moderate quality estimate of another species, but potential difference in detectability between species &  &  & 31 &  &  & few tens \\ 
   \hline
American pipit & mean & Derived from the abundance estimate of Lapland longspur using transects observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 53 &  &  & tens to few hundreds \\ 
   \hline
American pipit & mean & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 87 &  &  & tens to few hundreds \\ 
   \hline
snow bunting & mean & Derived from the abundance estimate of Lapland longspur using transects observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 18 &  &  & few hundreds \\ 
   \hline
snow bunting & mean & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 276 &  &  & few hundreds \\ 
   \hline
Arctic hare & mean & Derived from the abundance estimate of Arctic fox using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 6 &  &  & few tens \\ 
   \hline
\hline
\label{table:summary_methods}
\end{longtable}
\endgroup