### 2024-10-01 MetadataS1/tables/table_species_name_strategy.tex

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Tue Oct  1 09:17:32 2024
\begin{table}[ht]
\centering
\caption{Species composition of the vertebrate community of Bylot Island with the corresponding migratory status (i.e., resident, partial migrant or migrant).} 
\label{table:table_species_name_strategy}
\begingroup\fontsize{10pt}{10pt}\selectfont
\begin{tabularx}{0.95\textwidth}{llll}
  \hline
Functional group & Scientific name & English name & Migratory status \\ 
  \hline
Ducks and loons  & Gavia pacifica & Pacific loon & migrant \\ 
  Ducks and loons  & Gavia stellata & Red-throated loon & migrant \\ 
  Ducks and loons  & Somateria spectabilis & King eider & migrant \\ 
  Ducks and loons  & Clangula hyemalis & Long-tailed duck & migrant \\ 
  Geese and swans  & Branta hutchinsii & Cackling goose & migrant \\ 
  Geese and swans  & Anser caerulescens & Snow goose & migrant \\ 
  Geese and swans  & Cygnus columbianus & Tundra swan & migrant \\ 
  Raptors  & Buteo lagopus & Rough-legged hawk & migrant \\ 
  Raptors  & Falco peregrinus & Peregrine falcon & migrant \\ 
  Raptors  & Bubo scandiacus & Snowy owl & migrant \\ 
  Ptarmigans  & Lagopus muta & Rock ptarmigan & resident \\ 
  Cranes  & Antigone canadensis & Sandhill crane & migrant \\ 
  Shorebirds  & Pluvialis dominica & American golden-plover & migrant \\ 
  Shorebirds  & Pluvialis squatarola & Black-bellied plover & migrant \\ 
  Shorebirds  & Charadrius hiaticula & Common Ringed plover & migrant \\ 
  Shorebirds  & Arenaria interpres & Ruddy turnstone & migrant \\ 
  Shorebirds  & Calidris canutus & Red knot & migrant \\ 
  Shorebirds  & Calidris melanotos & Pectoral sandpiper & migrant \\ 
  Shorebirds  & Calidris bairdii & Baird's sandpiper & migrant \\ 
  Shorebirds  & Calidris fuscicollis & White-rumped sandpiper & migrant \\ 
  Shorebirds  & Calidris subruficollis & Buff-breasted sandpiper & migrant \\ 
  Shorebirds  & Phalaropus fulicarius & Red phalarope & migrant \\ 
  Gulls  & Larus hyperboreus & Glaucous gull & migrant \\ 
  Jaegers  & Stercorarius longicaudus & Long-tailed jaeger & migrant \\ 
  Jaegers  & Stercorarius parasiticus & Parasitic jaeger & migrant \\ 
  Ravens  & Corvus corax & Common raven & partial migrant \\ 
  Passerines & Eremophila alpestris & Horned lark & migrant \\ 
  Passerines & Anthus rubescens & American pipit & migrant \\ 
  Passerines & Calcarius lapponicus & Lapland longspur & migrant \\ 
  Passerines & Plectrophenax nivalis & Snow bunting & migrant \\ 
  Lemmings  & Lemmus trimucronatus & Nearctic brown lemming & resident \\ 
  Lemmings  & Dicrostonyx groenlandicus & Nearctic collared lemming & resident \\ 
  Hares  & Lepus arcticus & Arctic hare & resident \\ 
  Ermines  & Mustela richardsonii & Ermine & resident \\ 
  Foxes  & Vulpes lagopus & Arctic fox & partial migrant \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}