### 2024-10-01 MetadataS1/tables/table_species_year_monitoring.tex

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Tue Oct  1 09:17:34 2024
\begin{table}[ht]
\centering
\caption{Summary of vertebrate species monitoring in the Bylot Island study area. In this paper, we excluded certain years for specific species due to reduced sampling efforts. As a result, duration of times series presented here may differ slightly from those in Gauthier et al. (2024b).} 
\label{table:table_species_year_monitoring}
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{tabularx}{\textwidth}{lllll}
  \hline
Species & Zone & Years & Number of years & Monitoring \\ 
  \hline
Snow goose & camp 2 & 1999-2019, 2023 & (22) & systematic \\ 
  Rough-legged hawk & qarlikt., black \& south plat. & 2007-2019, 2022 & (15) & systematic \\ 
  Peregrine falcon & qarlikt., black \& south plat. & 2007-2019, 2022 & (15) & systematic \\ 
  Snowy owl & qarlikt., black \& south plat. & 1996-2019, 2023 & (25) & systematic \\ 
  Baird's sandpiper & qarlikturvik (2x1 km plot) & 2005-2019, 2022-2023 & (17) & systematic \\ 
  Lapland longspur & qarlikturvik (2x1 km plot) & 2005-2019, 2022-2023 & (17) & systematic \\ 
  King eider & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
  Long-tailed duck & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
  Rock ptarmigan & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
  Sandhill crane & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
  Nearctic brown lemming & qarlikturvik (trapping grids) & 1995-2019, 2021-2022 & (27) & systematic \\ 
  Nearctic collared lemming & qarlikturvik (trapping grids) & 1995-2019, 2021-2022 & (27) & systematic \\ 
  Pacific loon & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Red-throated loon & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Cackling goose & qarlikturvik valley & 2004-2019, 2022-2023 & (18) & systematic \\ 
  Tundra swan & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Glaucous gull & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Long-tailed jaeger & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Parasitic jaeger & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Pacific loon & whole study area & 2017-2019, 2022 & (4) & systematic \\ 
  Red-throated loon & whole study area & 2017-2019, 2022 & (4) & systematic \\ 
  Cackling goose & whole study area & 2017-2019, 2022-2023 & (5) & systematic \\ 
  Tundra swan & whole study area & 2017-2019, 2022 & (4) & systematic \\ 
  Rough-legged hawk & whole study area & 2013-2019, 2022 & (8) & systematic \\ 
  Peregrine falcon & whole study area & 2013-2019, 2022  & (8) & systematic \\ 
  Snowy owl & whole study area & 2012-2019, 2022-2023 & (10) & systematic \\ 
  Common Ringed plover & whole study area & 2015-2017 & (3) & systematic \\ 
  Glaucous gull & whole study area & 2017-2019, 2022 & (4) & systematic \\ 
  Parasitic jaeger & whole study area & 2009-2019, 2022 & (12) & opportunistic \\ 
  Common raven & whole study area & 2013-2019, 2022 & (8) & systematic \\ 
  Ermine & whole study area & 1993-2019 & (27) & opportunistic \\ 
  Arctic fox & whole study area & 2008-2016 & (9) & systematic \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}