### 2024-10-03 MetadataS1/tables/table_species_name_strategy.tex
Revision of the entire document and minor changes

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Thu Oct  3 11:06:53 2024
\begin{table}[ht]
\centering
\caption{Species of the vertebrate community of Bylot Island and their corresponding migratory status (i.e., resident, partial migrant or migrant).} 
\label{table:table_species_name_strategy}
\begingroup\fontsize{10pt}{10pt}\selectfont
\begin{tabularx}{0.97\textwidth}{llll}
  \hline
{\textbf{Functional group}} & {\textbf{Scientific name}} & {\textbf{English name}} & {\textbf{Migratory status}} \\ 
  \hline
Ducks and loons  & {\emph{ Gavia pacifica}} & Pacific loon & migrant \\ 
  Ducks and loons  & {\emph{ Gavia stellata}} & Red-throated loon & migrant \\ 
  Ducks and loons  & {\emph{ Somateria spectabilis}} & King eider & migrant \\ 
  Ducks and loons  & {\emph{ Clangula hyemalis}} & Long-tailed duck & migrant \\ 
  Geese and swans  & {\emph{ Branta hutchinsii}} & Cackling goose & migrant \\ 
  Geese and swans  & {\emph{ Anser caerulescens}} & Snow goose & migrant \\ 
  Geese and swans  & {\emph{ Cygnus columbianus}} & Tundra swan & migrant \\ 
  Raptors  & {\emph{ Buteo lagopus}} & Rough-legged hawk & migrant \\ 
  Raptors  & {\emph{ Falco peregrinus}} & Peregrine falcon & migrant \\ 
  Raptors  & {\emph{ Bubo scandiacus}} & Snowy owl & migrant \\ 
  Ptarmigans  & {\emph{ Lagopus muta}} & Rock ptarmigan & resident \\ 
  Cranes  & {\emph{ Antigone canadensis}} & Sandhill crane & migrant \\ 
  Shorebirds  & {\emph{ Pluvialis dominica}} & American golden-plover & migrant \\ 
  Shorebirds  & {\emph{ Pluvialis squatarola}} & Black-bellied plover & migrant \\ 
  Shorebirds  & {\emph{ Charadrius hiaticula}} & Common-ringed plover & migrant \\ 
  Shorebirds  & {\emph{ Arenaria interpres}} & Ruddy turnstone & migrant \\ 
  Shorebirds  & {\emph{ Calidris canutus}} & Red knot & migrant \\ 
  Shorebirds  & {\emph{ Calidris melanotos}} & Pectoral sandpiper & migrant \\ 
  Shorebirds  & {\emph{ Calidris bairdii}} & Baird's sandpiper & migrant \\ 
  Shorebirds  & {\emph{ Calidris fuscicollis}} & White-rumped sandpiper & migrant \\ 
  Shorebirds  & {\emph{ Calidris subruficollis}} & Buff-breasted sandpiper & migrant \\ 
  Shorebirds  & {\emph{ Phalaropus fulicarius}} & Red phalarope & migrant \\ 
  Gulls  & {\emph{ Larus hyperboreus}} & Glaucous gull & migrant \\ 
  Jaegers  & {\emph{ Stercorarius longicaudus}} & Long-tailed jaeger & migrant \\ 
  Jaegers  & {\emph{ Stercorarius parasiticus}} & Parasitic jaeger & migrant \\ 
  Ravens  & {\emph{ Corvus corax}} & Common raven & partial migrant \\ 
  Passerines & {\emph{ Eremophila alpestris}} & Horned lark & migrant \\ 
  Passerines & {\emph{ Anthus rubescens}} & American pipit & migrant \\ 
  Passerines & {\emph{ Calcarius lapponicus}} & Lapland longspur & migrant \\ 
  Passerines & {\emph{ Plectrophenax nivalis}} & Snow bunting & migrant \\ 
  Lemmings  & {\emph{ Lemmus trimucronatus}} & Nearctic brown lemming & resident \\ 
  Lemmings  & {\emph{ Dicrostonyx groenlandicus}} & Nearctic collared lemming & resident \\ 
  Hares  & {\emph{ Lepus arcticus}} & Arctic hare & resident \\ 
  Ermines  & {\emph{ Mustela richardsonii}} & Ermine & resident \\ 
  Foxes  & {\emph{ Vulpes lagopus}} & Arctic fox & partial migrant \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}