### 2024-10-03 MetadataS1/tables/table_species_year_monitoring.tex

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Thu Oct  3 13:54:34 2024
\begin{table}[ht]
\centering
\caption{Summary of vertebrate species monitoring in the Bylot Island study area. In this paper, we excluded certain years for specific species due to reduced sampling efforts. As a result, duration of times series presented here may differ slightly from those in Gauthier et al. (2024b).} 
\label{table:table_species_year_monitoring}
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{tabularx}{\textwidth}{lllll}
  \hline
{\textbf{Species}} & {\textbf{Zone}} & {\textbf{Years}} & {\textbf{Number of years}} & {\textbf{Monitoring}} \\ 
  \hline
Pacific loon & Qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Pacific loon & Whole study area & 2017-2019, 2022 & (4) & systematic \\ 
  Red-throated loon & Qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Red-throated loon & Whole study area & 2017-2019, 2022 & (4) & systematic \\ 
  King eider & Qarlikturvik (8 km2 plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
  Long-tailed duck & Qarlikturvik (8 km2 plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
  Cackling goose & Qarlikturvik valley & 2004-2019, 2022-2023 & (18) & systematic \\ 
  Cackling goose & Whole study area & 2017-2019, 2022-2023 & (5) & systematic \\ 
  Snow goose & Camp 2 & 1999-2019, 2023 & (22) & systematic \\ 
  Tundra swan & Qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Tundra swan & Whole study area & 2017-2019, 2022 & (4) & systematic \\ 
  Rough-legged hawk & Qarlikt., Black \& South plat. & 2007-2019, 2022 & (15) & systematic \\ 
  Rough-legged hawk & Whole study area & 2013-2019, 2022 & (8) & systematic \\ 
  Peregrine falcon & Qarlikt., Black \& South plat. & 2007-2019, 2022 & (15) & systematic \\ 
  Peregrine falcon & Whole study area & 2013-2019, 2022  & (8) & systematic \\ 
  Snowy owl & Qarlikt., Black \& South plat. & 1996-2019, 2023 & (25) & systematic \\ 
  Snowy owl & Whole study area & 2012-2019, 2022-2023 & (10) & systematic \\ 
  Rock ptarmigan & Qarlikturvik (8 km2 plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
  Sandhill crane & Qarlikturvik (8 km2 plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
  Common-ringed plover & Whole study area & 2015-2017 & (3) & systematic \\ 
  Baird's sandpiper & Qarlikturvik (2 km2 plot) & 2005-2019, 2022-2023 & (17) & systematic \\ 
  Glaucous gull & Qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Glaucous gull & Whole study area & 2017-2019, 2022 & (4) & systematic \\ 
  Long-tailed jaeger & South Qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Parasitic jaeger & South Qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Parasitic jaeger & Whole study area & 2009-2019, 2022 & (12) & opportunistic \\ 
  Common raven & Whole study area & 2013-2019, 2022 & (8) & systematic \\ 
  Lapland longspur & Qarlikturvik (2 km2 plot) & 2005-2019, 2022-2023 & (17) & systematic \\ 
  Nearctic brown lemming & Qarlikturvik (trapping grids) & 1995-2019, 2021-2022 & (27) & systematic \\ 
  Nearctic collared lemming & Qarlikturvik (trapping grids) & 1995-2019, 2021-2022 & (27) & systematic \\ 
  Ermine & Whole study area & 1993-2019 & (27) & opportunistic \\ 
  Arctic fox & Whole study area & 2008-2016 & (9) & systematic \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}