### 2024-11-08 MetadataS1/tables/interannual_nest_density_variation.tex
update manuscrit data paper open research statement and minor changes lapland longspur and Baird sandpiper

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Wed Nov  6 17:11:23 2024
\begin{table}[ht]
\centering
\caption{Annual nest density (nests/km2) of selected avian species estimated on different zones of Bylot Island.} 
\label{table:interannual_nest_density_variation}
\begingroup\fontsize{10pt}{10pt}\selectfont
\begin{tabularx}{0.9\textwidth}{lllr}
  \hline
Species & Zone & Mean ± SD & Number of years \\ 
  \hline
Pacific loon & Whole study area & 0.005 ± 0.004 &   4 \\ 
  Red-throated loon & Whole study area & 0.082 ± 0.019 &   4 \\ 
  King eider & Qarlikturvik (4x2 km plot) & 0.115 ± 0.138 &  16 \\ 
  Long-tailed duck & Qarlikturvik (4x2 km plot) & 0.092 ± 0.138 &  16 \\ 
  Cackling goose & Whole study area & 0.177 ± 0.064 &   5 \\ 
  Tundra swan & Whole study area & 0.001 ± 0.001 &   4 \\ 
  Rough-legged hawk & Whole study area & 0.157 ± 0.151 &   8 \\ 
  Peregrine falcon & Whole study area & 0.053 ± 0.007 &   8 \\ 
  Snowy owl & Whole study area & 0.022 ± 0.058 &  10 \\ 
  Rock ptarmigan & Qarlikturvik (4x2 km plot) & 0.031 ± 0.055 &  16 \\ 
  Sandhill crane & Qarlikturvik valley & 0.088 ± 0.055 &  17 \\ 
  Common-ringed plover & Whole study area & 0.070 ± 0.012 &   3 \\ 
  Baird's sandpiper & Qarlikturvik (2x1 km plot) & 5.000 ± 3.558 &  17 \\ 
  Glaucous gull & Whole study area & 0.091 ± 0.011 &   4 \\ 
  Long-tailed jaeger & Qarlikturvik valley & 0.362 ± 0.380 &  17 \\ 
  Parasitic jaeger & Whole study area & 0.010 ± 0.004 &  12 \\ 
  Common raven & Whole study area & 0.003 ± 0.003 &   8 \\ 
  Lapland longspur & Qarlikturvik (2x1 km plot) & 13.559 ± 5.849 &  17 \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}